Movie Recommendation System

### 2021-03-17 draft3.Rmd
updated the draft and finished the main

Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`

To improve this model, it is possible to account for the fact that movies are not received by audiences equally. A bias term is calculated by taking the average of the difference between the individual ratings of a movie and the average of all movie ratings. This will represent how a particular movie deviates from the average movie, and will be added onto our previous model: 
```{r, echo=FALSE, message=FALSE}
# This model builds on previous by introducing a movie bias term
movie_avgs <- edx_train %>% group_by(movieId) %>% summarise(b_i=mean(rating-mu))
movie_bias_model <- mu + edx_test %>% left_join(movie_avgs, by='movieId') %>% pull(b_i)
movie_bias_rmse <- RMSE(edx_test$rating, movie_bias_model)
```
Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`
Avg + Movie Bias                          | `movie_bias_rmse`

Likewise, different users have different movie critiquing tendencies. A bias term for users, is calculated by taking the average of each users rating after subtracting the average movie rating and movie bias term. The user bias term will account for how a particular movie deviates from the average movie for a particular user. Again this term is added onto the previous model:
```{r, echo=FALSE, message=FALSE}
# This model builds on previous by introducing a user bias term
user_avgs <- edx_train %>% left_join(movie_avgs, by='movieId') %>% group_by(userId) %>% summarise(b_u=mean(rating-mu-b_i))
user_movie_bias_model <- edx_test %>% left_join(movie_avgs, by='movieId') %>% left_join(user_avgs, by='userId') %>%
  mutate(pred=mu+b_i+b_u) %>% pull(pred)
user_movie_bias_rmse <- RMSE(edx_test$rating, user_movie_bias_model)
```
Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`
Avg + Movie Bias                          | `movie_bias_rmse`
Avg + Movie Bias + User Bias              | `user_movie_bias_rmse`

@@

# This model builds on previous by introducing a movie bias term
movie_avgs <- edx_train %>% left_join(user_avgs, by='userId') %>% group_by(movieId) %>% summarise(b_i=mean(rating-mu-b_u))
movie_user_bias_model <- edx_test %>% left_join(movie_avgs, by='movieId') %>% left_join(user_avgs, by='userId') %>%
  mutate(pred=mu+b_i+b_u) %>% pull(pred)
movie_user_bias_rmse <- RMSE(edx_test$rating, movie_user_bias_model)
```
Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`
Avg + Movie Bias                          | `movie_bias_rmse`
Avg + Movie Bias + User Bias              | `user_movie_bias_rmse`
Avg + User Bias                           | `user_bias_rmse`
Avg + User Bias + Movie Bias              | `movie_user_bias_rmse`

@@

```
Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`
Avg + Movie Bias                          | `movie_bias_rmse`
Avg + Movie Bias + User Bias              | `user_movie_bias_rmse`
Avg + User Bias + Movie Bias              | `movie_user_bias_rmse`
Avg + Movie Bias + User Bias + Lambda     | `regularized_user_movie_rmse`

@@

Model                                     |    RMSE
------------------------------------------|-------------
Average                                   | `naive_rmse`
Avg + Movie Bias                          | `movie_bias_rmse`
Avg + Movie Bias + User Bias              | `user_movie_bias_rmse`
Avg + User Bias + Movie Bias              | `movie_user_bias_rmse`
Avg + Movie Bias + User Bias + Lambda     | `regularized_user_movie_rmse`
Matrix Factorization (Untuned)            | `matrix_factor_rmse`
Matrix Factorization (Tuned)              | `matrix_factor_tuned_rmse`

#Results
###section that presents the modeling results and discusses the model performance
We now have several models, and we will select the best performing one We will now run our best performing model on the validation set
```{r, echo=FALSE}

#Conclusion
###section that gives a brief summary of the report, its limitations and future work

>>>>

To improve this model, it is possible to account for the fact that movies are not received by audiences equally. A bias term is calculated by taking the average of the difference between the individual ratings of a movie and the average of all movie ratings. This will represent how a particular movie deviates from the average movie, and will be added onto our previous model: 
```{r, echo=FALSE, message=FALSE}
# This model builds on previous by introducing a movie bias term
movie_avgs <- edx_train %>% group_by(movieId) %>% summarise(b_i=mean(rating-mu))
movie_bias_model <- mu + edx_test %>% left_join(movie_avgs, by='movieId') %>% pull(b_i)
movie_bias_rmse <- RMSE(edx_test$rating, movie_bias_model)
```

Likewise, different users have different movie critiquing tendencies. A bias term for users, is calculated by taking the average of each users rating after subtracting the average movie rating and movie bias term. The user bias term will account for how a particular movie deviates from the average movie for a particular user. Again this term is added onto the previous model:
```{r, echo=FALSE, message=FALSE}
# This model builds on previous by introducing a user bias term
user_avgs <- edx_train %>% left_join(movie_avgs, by='movieId') %>% group_by(userId) %>% summarise(b_u=mean(rating-mu-b_i))
user_movie_bias_model <- edx_test %>% left_join(movie_avgs, by='movieId') %>% left_join(user_avgs, by='userId') %>%
  mutate(pred=mu+b_i+b_u) %>% pull(pred)
user_movie_bias_rmse <- RMSE(edx_test$rating, user_movie_bias_model)
```

@@

# This model builds on previous by introducing a movie bias term
movie_avgs <- edx_train %>% left_join(user_avgs, by='userId') %>% group_by(movieId) %>% summarise(b_i=mean(rating-mu-b_u))
movie_user_bias_model <- edx_test %>% left_join(movie_avgs, by='movieId') %>% left_join(user_avgs, by='userId') %>%
  mutate(pred=mu+b_i+b_u) %>% pull(pred)
movie_user_bias_rmse <- RMSE(edx_test$rating, movie_user_bias_model)
```

@@

@@

This method of matrix factorization gives the best result. Even without tuning the parameters, it beats the other models.

# Results
### section that presents the modeling results and discusses the model performance

The best performing model is selected to use on the validation set. It is trained with the full edx data set and the rmse of the predictions is calculated below:

```{r}
```

# Conclusion
###section that gives a brief summary of the report, its limitations and future work

In practice the recommendation system would only recommend movies that are predicted to be above the users average rating. This validation set replicates how future users would rate movies, both the movies that were recommended and movies that were not recommended (due to a low score).

As the data set is constantly growing, and the algorithm is being tested in the wild, it is realistic to check the rmse of the model post-implementation. A data pipeline of importing the newest results and adding them to the existing model would also be important. Verifying that the updating rmse is behaving as expected is an easy way of making sure the model has not broken down.
